From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-29 Pi+/paper/finite.tex
Fix the proposition about cox equiv longcox

\begin{definition}
  Let $A$ be an $\hSet$, and $\List[\blank]$ the free monoid. The free group
  $F(A)$ on $A$ is the set-quotient of $\List[A + A]$ by the congruence closure
  of the relation $a \cons \inv{a} \cons \nil \sim \nil$.
\end{definition}

@@

The type $\Sn$ is then defined as the set-quotient of $\List[\Fin[n]]$ by $\cox*$.

@@

Fortunately, the change from the rewriting system described before will be small. Thus, we will not lose the connection to the theory of group presentations, where Coxeter relations are a standard notion~\todo{citation} -- to make the result broadly applicable, even though we use the reduction system for computation, we prove that its reflexive-transitive closure is indeed equivalent to the conguence closure of the standard Coxeter relation $\cox$.

@@

First, we need a helper function $n \downf k$.

@@

The reuslt of this function is the sequence \(k + n, k + n - 1, k + n - 2, \ldots, n + 1\). If we think of the element $m$ as the transposition which swaps elements $\el{m}$ and $\el{m+1}$, the role of this helper function is to produce a sequence of transpositions - a permutation - which moves element $n$ $k$ places right.

As a sequence of transpositions, it moves the element $k - 1$ places. Expressed in terms of the braiding diagram, for
$n = \dots$ and $k = \dots$, it has the following form:

\[
  \begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd,
      number of strands = 6,
      thick]
    {braid={s_5, s_4, s_3, s_2, s_1}};
  \end{tikzpicture}
\]

Formally, the directed $\longcox$ relation is defined with the following generators.

@@

Like before, we define the $\longcox*$ to be the closure of $\longcox$, but only under reflexivity and transitivity. The
transitive extension is only performed on the right.

@@

Generators $\longcancel$ and $\longswap$ correspond directly to the generators of $\cox*$ and can be visualised in the
same way as before. The generator $\longbraid$

Now, the remaining constructor $\longbraid$ moves a transposition around such sequence, as in the diagram below.

Note that the previous $\braid$ rule is a special case of $\longbraid$, with $k = 2$.

@@

The modified form of the Coxeter relations are unwieldy and difficult to prove properties about by induction, and have
little connection to the broader scope of computational group theory. However, we can prove that the closures of the two
forms of the relations are equivalent, and work up to this equivalence!

\begin{proposition}
  $\cox*$ and $\longcox*$ are equivalent. For every $w$ and $v$, $w \cox* v$ iff $w \longcox* v$.
\end{proposition}

>>>>

\begin{definition}~\label{def:presentation}
  Let $A$ be an $\hSet$, and $\List[\blank]$ the free monoid. The free group
  $F(A)$ on $A$ is the set-quotient of $\List[A + A]$ by the congruence closure
  of the relation $a \cons \inv{a} \cons \nil \sim \nil$.
\end{definition}

@@

\jk{maybe change that?}
Group persentation, as given in the ~\label{def:presentation}, requires the relation to be on the set of $A + A$, where the right copy corresponded to the set of formal inverses of the generators. However, in our case, we can observe that that the constructor $\cancel$ specifies that the inverse of each element is again the same element. Thus, the type $\Sn$ is defined as the set-quotient of $\List[\Fin[n]]$ by $\cox*$.

@@

Fortunately, the change from the rewriting system described before will be small. Thus, we will not lose the connection to the theory of group presentations, where Coxeter relations are a standard notion~\todo{citation} -- to make the result broadly applicable, even though we use the reduction system for computation, we prove that the \jk{relation it gives} is indeed equivalent to the conguence closure of the standard Coxeter relation $\cox$.

@@

First, we need to define a helper function $n \downf k$.

@@

The result of this function is the sequence \(k + n, k + n - 1, k + n - 2, \ldots, n + 1\). Since we think of the number $m$ as the transposition $\tau_m$ which swaps elements $\el{m}$ and $\el{m+1}$, the role of this helper function is to produce a sequence of transpositions -- a permutation -- which moves element $\el{k + n}$ by $k$ places left, shifting all the elements in between one place right. Expressed in terms of the braiding diagram, for
$n = 0$ and $k = 4$, it has the following form:

\[
  \begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd,
      number of strands = 5,
      thick]
    {braid={s_4, s_3, s_2, s_1}};
  \end{tikzpicture}
\]

Then, the directed relation $\longcox$ is defined with the following generators.

@@

Constructors $\longcancel$ and $\longswap$ correspond directly to the approproiate constructors of $\cox$ and can be visualised in the same way as before. The remaining constructor uses the helper function to exchange the order of a long sequence of transpositions and a single transposition afterwards. For example, for $n = 2$ and $k = 4$, it allows for the reduction
$[7, 6, 5, 4, 3, 7] \longcox* [6, 7, 6, 5, 4, 3]$. Visualized on the braid diagram, it looks as follows:

Note that the previous $\braid$ rule is a special case of $\longbraid$, with $k = 0$.

Like before, we define the $\longcox*$ to be the closure of $\longcox$, but only under reflexivity and transitivity.

@@

@@

The modified form of the Coxeter relations are unwieldy and difficult to prove properties about by induction, and, as we mentioned, have little connection to the broader scope of computational group theory. However, we can prove that the closures of the two forms of the relations are equivalent, and work up to this equivalence!

\begin{proposition}~\label{prop:coxlongcox}
  $\cox*$ and $\longcox*$ are equivalent in the following sense: for every $w$ and $v$, $w \cox* v$ iff there is a $u$ such that $w \longcox* u$ and $v \longcox* u$.
\end{proposition}
\jk{does the construction above has a name? It is analogous to the one we do in lambda calculus, when taking $\equiv_{\beta}$. We have to write something about this proof using the Church-Rosser idea.}
